---
title: "CARE DREAMS - Malawi"
author: "Mia Sumera"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    orientation: rows
---

```{r setup, include=FALSE}
library(flexdashboard)
library(here)
library(tidyverse)
library(DT)
library(highcharter)
library(knitr)

here::i_am("DREAMS_Dashboard.Rmd")

samples      <- readRDS(here("output/samples.rds"))
regressions  <- readRDS(here("output/regressions.rds"))
sum_data     <- readRDS(here("output/summary_data.rds"))
figs         <- readRDS(here("output/figures.rds"))
reg_table    <- readRDS(here("output/reg_table.rds"))
care_map     <- readRDS(here("output/care_map.rds"))
assist_table <- readRDS(here("output/assist_table.rds"))

fmt_p <- regressions$fmt_p

make_rcdf <- function(score_vec, group_label, max_score = 8) {
  counts <- table(factor(floor(score_vec), levels = 0:max_score))
  total  <- sum(counts)
  tibble(
    score = 0:max_score,
    pct   = map_dbl(0:max_score,
                    ~ sum(counts[(.x + 1):length(counts)]) / total * 100),
    group = group_label
  )
}
```

Overview 
=====================================

Column {.sidebar}
-------------------------------------

**Diversifying Resilience Approaches through Market Systems in Emergency (DREAMS)**

**Introduction**

The 2023/2024 El Niño event produced severe food and nutrition insecurity across southern Malawi. In response, the DREAMS program was implemented across Chiradzulu, Phalombe, and Thyolo through four sectors:

- Food Assistance
- Agriculture Support
- Nutrition Services
- Protection Programming

**Research Question**

How does the integrated delivery of unconditional food assistance (UFA) and supplemental nutrition assistance (SNA) top-ups influence nutritional status for children under two and women, compared to households receiving UFA alone?

**Study Population** 

The analytic sample includes Women of Reproductive Age (WRA, 15-49) and households with children under two (6–23months) within DREAMS beneficiary households. 

---

**Note:** *Households that received Unconditional Food Assistance also received Unconditional Cash Transfer.* 

Row {data-height=200}
-------------------------------------
### Assistance Coverage
```{r table}
assist_table
```

Row
-------------------------------------
### Program Districts — Malawi
```{r map}
care_map
```

### Assistance Distribution
```{r assist-bar}
figs$assist_bar
```


Food Consumption Score
=====================================

Column {.sidebar}
-------------------------------------

**Food Consumption Score (FCS)** measures the dietary quality and diversity of a household based on how often they consumed different food groups in the past 7 days.

**UN World Food Programme (WFP) Thresholds:**

- **Acceptable:** FCS ≥ 35
- **Borderline:** FCS 21.5–35
- **Poor:** FCS < 21

**Model:** Adjusted linear regression. β = mean difference in FCS for UFA+SNA vs UFA only, holding all covariates constant.

**Covariates:** Age, marital status, education, household status, household size, sex of HH head, district.

---

**Key Result:**

```{r fcs-key}
beta_val <- round(regressions$fcs_int$estimate, 2)
htmltools::p(
  style = "font-size:13px;",
  htmltools::strong("\u03b2 = "), beta_val,
  htmltools::br(),
  htmltools::strong("95% CI: "),
  paste0("[", round(regressions$fcs_int$ci_lower, 2), ", ",
         round(regressions$fcs_int$ci_upper, 2), "]"),
  htmltools::br(),
  htmltools::strong("p = "), fmt_p(regressions$fcs_int$p.value)
)
```


Row
-------------------------------------

### FCS Category Distribution by Program Arm
```{r fcs-cat}
figs$fcs_cat
```

Row {data-height=210}
-------------------------------------

### Crude & Adjusted Effect Estimates {data-width=500}
```{r reg-table-fcs}
reg_table$reg_table_fcs
```


### Key Takeaway {data-width=500}
```{r fcs-takeaway}
htmltools::div(
  style = "padding: 8px 12px; font-family: Arial; font-size: 13px; line-height: 1.7;",
  htmltools::p(
    htmltools::strong("UFA+SNA functioned as a protective factor"),
    " against seasonal food insecurity: while UFA-only households saw FCS decline at endline
    (e.g. Thyolo: 35.3 \u2192 26.3), UFA+SNA maintained near-baseline levels (35.1), ",
    htmltools::strong("scored significantly higher"),
    " after adjustment (",
    htmltools::strong(paste0("\u03b2 = ", round(regressions$fcs_int$estimate, 2),
                             ", p = ", fmt_p(regressions$fcs_int$p.value))),
    "), and had less than half the proportion of food-insecure households (10.7% vs 26.7% Poor)."
  ),
  htmltools::p(
    "Integrated assistance yielded the lowest percentage of households in the Poor category
    (FCS < 21) and the highest in both Borderline and Acceptable categories, suggesting
    the nutrition top-up provides ",
    htmltools::strong("greater protection"),
    " against food insecurity beyond food assistance alone."
  )
)
```


Minimum Diet Diversity - Women
=====================================

Column {.sidebar}
-------------------------------------

**Minimum Dietary Diversity — Women (MDD-W)** measures whether women aged 15–49 consumed food from ≥5 of 8 food groups in the past 24 hours.

**MDD threshold:**

≥ 5 food groups = MDD **met**

< 5 food groups = MDD **not met**


**Model:** Adjusted logistic regression. Sex of household head excluded (female-only subgroup).

**Covariates:** Age, marital status, education, household status, household size, district.

---

**Key Result:**

```{r mddw-key}
htmltools::p(
  style = "font-size:13px;",
  htmltools::strong("OR = "), round(regressions$mddw_int$OR, 2),
  htmltools::br(),
  htmltools::strong("95% CI: "),
  paste0("[", round(regressions$mddw_int$ci_lower, 2), ", ",
         round(regressions$mddw_int$ci_upper, 2), "]"),
  htmltools::br(),
  htmltools::strong("p = "), fmt_p(regressions$mddw_int$p.value)
)
```

Row
-------------------------------------

### Women (15–49 years) — Reverse Cumulative Distribution
```{r cdf-mddw}
cdf_cols <- c("Baseline" = "#aaaaaa55", "Unconditional Food Assistance (UFA only)" = "#5F5E5A80", "Unconditional Food Assistance + Nutrition Top-up (UFA+SNA)" = "#C45E0A")

cdf_w <- bind_rows(
  make_rcdf(samples$mddw_bl$mdd_w_score,  "Baseline"),
  make_rcdf(samples$mddw_ufa$mdd_w_score, "Unconditional Food Assistance (UFA only)"),
  make_rcdf(samples$mddw_sna$mdd_w_score, "Unconditional Food Assistance + Nutrition Top-up (UFA+SNA)")
) %>%
  mutate(group = factor(group, levels = c("Baseline", 
                                          "Unconditional Food Assistance (UFA only)", 
                                          "Unconditional Food Assistance + Nutrition Top-up (UFA+SNA)")),
         pct = round(pct, 1))

hchart(cdf_w, "line", hcaes(x = score, y = pct, group = group)) %>%
  hc_colors(unname(cdf_cols)) %>%
  hc_title(text = "MDD-W: Minimum Dietary Diversity \u2014 Women (15\u201349)") %>%
  hc_subtitle(text = "Adjusted OR = 1.12 (p = 0.75) \u2014 not significant") %>%
  hc_xAxis(title = list(text = "Number of Food Groups Consumed (\u22655)"),
           min = 0, max = 8, tickInterval = 1,
           plotLines = list(list(value = 5, color = "#333333", dashStyle = "Dash", width = 2,
                                 label = list(text = "MDD threshold (\u22655)",
                                              style = list(color = "#333333", fontSize = "11px"))))) %>%
  hc_yAxis(title = list(text = "% of Women"), min = 0, max = 100,
           labels = list(format = "{value}%")) %>%
  hc_tooltip(useHTML = TRUE, headerFormat = "",
             pointFormat = "<b>Group:</b> {series.name}<br><b>Food groups \u2265:</b> {point.x}<br><b>Percentage:</b> {point.y:.1f}%") %>%
  hc_credits(enabled = TRUE,
             text = paste0("Baseline n=", nrow(samples$mddw_bl),
                           " | UFA only n=", nrow(samples$mddw_ufa),
                           " | UFA+SNA n=", nrow(samples$mddw_sna),
                           ". Sample: Households including WRA."),
             style = list(fontSize = "10px")) %>%
  hc_legend(enabled = TRUE, layout = "horizontal", align = "left", verticalAlign = "top") %>%
  hc_add_theme(hc_theme_google())
```

Row {data-height=160}
-------------------------------------

### Crude & Adjusted Effect Estimates {data-width=550}
```{r reg-table-w}
reg_table$reg_table_mddw
```


### Key Takeaway {data-width=450}

```{r mddw-takeaway}
htmltools::div(
  style = "padding: 8px 12px; font-family: Arial; font-size: 13px; line-height: 1.7;",
  htmltools::p(
    htmltools::strong("MDD-W trended positive"),
    " at endline for UFA+SNA vs UFA only (53.1% vs 45.5%) but did not
    reach statistical significance (OR = ",
    round(regressions$mddw_int$OR, 2), ", p = ",
    fmt_p(regressions$mddw_int$p.value),
    "), likely reflecting ",
    htmltools::strong("limited statistical power"),
    " (n = ", nrow(samples$mddw_sna),
    " in the UFA+SNA group)", htmltools::strong("rather than absence of effect.")
  ),
  htmltools::p(
    "Notably, only ",
    htmltools::strong("26.9% of mothers with infants aged 0\u20135 months met MDD-W at baseline"),
    ", a separate but related subgroup, reflecting high unmet nutritional need among the most
    vulnerable women, and underscoring that ",
    htmltools::strong(" food assistance alone is insufficient"),
    " to meet dietary diversity needs."
  )
)
```


Minimum Diet Diversity - Children
=====================================

Column {.sidebar}
-------------------------------------

**Minimum Dietary Diversity — Children (MDD-C)** measures whether children aged 6–23 months consumed food from ≥5 of 8 food groups in the past 24 hours.

**MDD threshold:**

≥ 5 food groups = MDD **met**

< 5 food groups = MDD **not met**


**Model:** Firth penalized logistic regression used due to small sample size and sparse cells in the UFA+SNA subgroup.

**Covariates:** Age, marital status, education, household status, household size, sex of household head, district.

---

**Key Result:**

```{r mddc-key}
htmltools::p(
  style = "font-size:13px;",
  htmltools::strong("OR = "), round(regressions$mddc_int$OR, 2),
  htmltools::br(),
  htmltools::strong("95% CI: "),
  paste0("[", round(regressions$mddc_int$ci_lower, 2), ", ",
         round(regressions$mddc_int$ci_upper, 2), "]"),
  htmltools::br(),
  htmltools::strong("p = "), fmt_p(regressions$mddc_int$p.value)
)
```

Row
-------------------------------------

### Children (6–23 months) — Reverse Cumulative Distribution
```{r cdf-mddc}
cdf_c <- bind_rows(
  make_rcdf(samples$mddc_bl$mdd_c_score,  "Baseline"),
  make_rcdf(samples$mddc_ufa$mdd_c_score, "Unconditional Food Assistance (UFA only)"),
  make_rcdf(samples$mddc_sna$mdd_c_score, "Unconditional Food Assistance + Nutrition Top-up (UFA+SNA)")
) %>%
  mutate(group = factor(group, levels = c("Baseline", 
                                          "Unconditional Food Assistance (UFA only)", 
                                          "Unconditional Food Assistance + Nutrition Top-up (UFA+SNA)")),
         pct = round(pct, 1))

hchart(cdf_c, "line", hcaes(x = score, y = pct, group = group)) %>%
  hc_colors(unname(cdf_cols)) %>%
  hc_title(text = "MDD-C: Minimum Dietary Diversity \u2014 Children (6\u201323m)") %>%
  hc_subtitle(text = "Firth adjusted OR = 1.29 (p = 0.52) \u2014 not significant") %>%
  hc_xAxis(title = list(text = "Number of Food Groups Consumed (\u2265)"),
           min = 0, max = 8, tickInterval = 1,
           plotLines = list(list(value = 5, color = "#333333", dashStyle = "Dash", width = 2,
                                 label = list(text = "MDD threshold (\u22655)",
                                              style = list(color = "#333333", fontSize = "11px"))))) %>%
  hc_yAxis(title = list(text = "% of Children"), min = 0, max = 100,
           labels = list(format = "{value}%")) %>%
  hc_tooltip(useHTML = TRUE, headerFormat = "",
             pointFormat = "<b>Group:</b> {series.name}<br><b>Food groups \u2265:</b> {point.x}<br><b>Percentage:</b> {point.y:.1f}%") %>%
  hc_credits(enabled = TRUE,
             text = paste0("Baseline n=", nrow(samples$mddc_bl),
                           " | UFA only n=", nrow(samples$mddc_ufa),
                           " | UFA+SNA n=", nrow(samples$mddc_sna),
                           ". Sample: Households including children (6\u201323m)."),
             style = list(fontSize = "10px")) %>%
  hc_legend(enabled = TRUE, layout = "horizontal", align = "left", verticalAlign = "top") %>%
  hc_add_theme(hc_theme_google())
```

Row {data-height=160}
-------------------------------------

### Crude & Adjusted Effect Estimates {data-width=550}

```{r reg-table-c}
reg_table$reg_table_mddc
```

### Key Takeaway {data-width=450}

```{r mddc-takeaway}
htmltools::div(
  style = "padding: 8px 12px; font-family: Arial; font-size: 13px; line-height: 1.7;",
  htmltools::p(
    htmltools::strong("MDD-C trended positive"),
    " at endline for UFA+SNA vs UFA only (",
    round(mean(samples$mddc_sna$mdd_c_cat, na.rm=TRUE)*100, 1), "% vs ",
    round(mean(samples$mddc_ufa$mdd_c_cat, na.rm=TRUE)*100, 1),
    "%) but did not reach statistical significance (OR = ",
    round(regressions$mddc_int$OR, 2), ", p = ",
    fmt_p(regressions$mddc_int$p.value),
    "), likely reflecting ",
    htmltools::strong("limited statistical power"),
    " (n = ", nrow(samples$mddc_sna),
    " in the UFA+SNA group) ",
    htmltools::strong("rather than absence of effect.")
  ),
  htmltools::p(
    "The consistent positive direction across all three outcomes (FCS, MDD-W, and MDD-C)
    suggests ", htmltools::strong("broader nutritional benefit of integrated assistance"),
    ". Results should be interpreted with caution given the small UFA+SNA child subsample
    and use of Firth penalized regression."
  )
)
```

About the Data
=====================================

Column {.sidebar}
-------------------------------------

**Data Access**

This dataset is **confidential** and not publicly available. Access is restricted to:

- CARE Malawi MEAL Team
- USAID/BHA
- CADECOM & GENET (programme partners)
- Malawi government counterparts

**Collected via:** KoboToolbox

**Analysed in:** R (flexdashboard, tidyverse, logistf, DT, highcharter)

Row
-------------------------------------

### Dataset Description

```{r about}
htmltools::div(
  style = "padding: 20px; font-family: Arial; font-size: 14px; line-height: 1.9; max-width: 900px;",

  htmltools::h4(style = "color:#C45E0A;", "Data Description"),
    htmltools::p(
      "This dashboard uses confidential household survey data collected by CARE Malawi's team via
      KoboToolbox at baseline (October 2024, post-harvest season) and
      endline (January 2026, lean season) across the Chiradzulu, Phalombe, and 
      Thyolo districts of southern Malawi."
    ),
    htmltools::p(
      "The study population comprised DREAMS program beneficiary households receiving
      at least one intervention component, with the analytic sample restricted to
      women of reproductive age (WRA, 15\u201349) and households with children aged
      6\u201323 months."
    ),
    htmltools::p(
      "Data were collected using structured household surveys
      administered by trained enumerators, capturing household food consumption,
      dietary diversity, assistance receipt, and demographic characteristics."
    ),
  
    htmltools::p(
      "Endline data are used for primary analysis, applying adjusted regression models
      to compare UFA+SNA versus UFA only households in a cross-sectional design;
     linear regression (\u03b2 coefficient) for FCS and logistic regression (odds ratio)
      for MDD-W and MDD-C, with Firth penalized logistic regression used for MDD-C
      due to sparse cells in the small UFA+SNA child subsample."
    ),

  htmltools::h4(style = "color:#C45E0A;", "Real-World Impact"),
  htmltools::p("With 149 million children under five affected by stunting and wasting globally,
  evidence on whether integrated nutrition assistance outperforms food assistance
  alone is urgently needed to guide humanitarian investment. These findings directly
  inform USAID/BHA funding decisions and CARE Malawi's program design by demonstrating
  that the SNA top-up component provides measurable protective benefits for food
  security and dietary diversity beyond unconditional food assistance alone."),

  htmltools::hr(style = "margin-top:20px; border-color:#ddd;"),
  htmltools::p(
    style = "font-size:11px; color:#888;",
    "DREAMS = Diversifying Resilience Approaches through Market Systems in Emergency . ",
    "UFA = Unconditional Food Assistance. SNA = Supplementary Nutrition Assistance. ",
    "FCS thresholds: Poor \u226421, Borderline 22\u201335, Acceptable >35 (WFP 2008). ",
    "MDD threshold: \u22655 food groups consumed in past 24 hours."
  )
)
```

